### 2024-07-05 chap_RL/chap_RL.tex
Rebuttals: Keep going...

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.49\textwidth]{Reward_status.pdf}
\includegraphics[width=0.49\textwidth]{Reward_status_2.pdf}
\caption{Reward distribution between successes and failures. Right hand side details at the end of which time-window the failure occurred. The ``tipping year'' is 2040 as failing the transition by 2040 represents 57\% of all the failures. Beyond this point, through this learning process, succeeding the transition represents 38\% of the episodes. }
\label{fig:reward_status}
\end{figure}

@@

The first two dimensions of the state space are the cumulative emissions and costs. They drive the value of the reward and, consequently, the optimisation of the agent's policy. Per definition, the threshold of 1.2\,Gt$_{\ce{CO2},\text{eq}}$ splits the episodes reaching 2050 into successes and failures (see Figure \ref{fig:Cum_gwp_cost}). In the successful transitions, the median cumulative emissions, $\text{P}_{50}$, are about 0.9\,Gt$_{\ce{CO2},\text{eq}}$. Reaching cumulative emissions significantly lower than the \ce{CO2}-budget are possible thanks to efforts made at earlier stages of the transition and the potential to install \gls{SMR} later on. Considering the failures, half of these episodes ended up with cumulative emissions lower or equal to 1.4\,Gt$_{\ce{CO2},\text{eq}}$. As illustrated in Figure \ref{fig:reward_status}, 2040 is identified as the tipping year. Where 98\% of the failures were below the \ce{CO2}-budget in 2035, only 37\% passed this threshold in 2040. This reminds the importance of near-term, 2025-2030, actions to hope for a successful transition.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Cum_gwp_cost_3.pdf}
\caption{Exploration of the state space over the learning process: cumulative emissions and costs. The majority of successful transitions have cumulative emissions much lower than the \ce{CO2}-budget and are cheaper than the REF case. }
\label{fig:Cum_gwp_cost}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{RE_in_mix_Efficiency.pdf}
\caption{Exploration of the state space over the learning process: share of renewable energy carriers in the primary energy mix and efficiency. Integration of local \gls{VRES} at early stages then massive import of electrofuels later are needed to secure successful transitions. Below a near-term threshold ($\sim$60\%), the chances of success are limited, \ie no-go zones. Efficiency is a less valuable information for the agent to succeed transitions as failures and successes indistinguishably spread over the whole range.}
\label{fig:RE_in_mix_Efficiency}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Actions_learning.pdf}
\caption{Besides the wide exploration of the action space, successes and failures indistinguishably spread over the whole ranges of actions. In other words, it does not seem to be any clear set of actions to support successful transitions.}
\label{fig:Actions_learning}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Binding_learning.pdf}
\caption{Depending on the action and its timing, it is actually constraining the optimisation through EnergyScope Pathway or not. Sweet spots can be identified when considering the limits of \gls{GWP} and fossil gas consumption. Limiting coal consumption is always constraining, unlike \gls{LFO} that is ``naturally'' substituted by EnergyScope Pathway in the near-term.}
\label{fig:Binding_learning}
\end{figure}

>>>>

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.49\textwidth]{Reward_status.pdf}
\includegraphics[width=0.49\textwidth]{Reward_status_2.pdf}
\caption{Reward distribution between successes and failures. Graph on the right hand side details at the end of which time-window the failure occurred. The ``tipping year'' is 2040 as failing the transition by 2040 represents 57\% of all the failures. Beyond this point, through this learning process, succeeding the transition represents 38\% of the episodes. }
\label{fig:reward_status}
\end{figure}

@@

The first two dimensions of the state space are the cumulative emissions and costs. They drive the value of the reward and, consequently, the optimisation of the agent's policy. Per definition, the threshold of 1.2\,Gt$_{\ce{CO2},\text{eq}}$ splits the episodes reaching 2050 into successes and failures (see Figure \ref{fig:Cum_gwp_cost}). Since infeasible cases or those that overshoot the \ce{CO2}-budget are discarded before reaching 2050 (see Section \ref{subsec:RL:act_states_rew:rew}), the number of attempts that reach further steps in the transition progressively decreases. Consequently, the share of successful transitions compared to failures progressively increases with time. In the successful transitions, the median cumulative emissions, $\text{P}_{50}$, are about 0.9\,Gt$_{\ce{CO2},\text{eq}}$. Reaching cumulative emissions significantly lower than the \ce{CO2}-budget are possible thanks to efforts made at earlier stages of the transition and the potential to install \gls{SMR} later on. Considering the failures, half of these episodes ended up with cumulative emissions lower or equal to 1.4\,Gt$_{\ce{CO2},\text{eq}}$. As illustrated in Figure \ref{fig:reward_status}, 2040 is identified as the tipping year. Where 98\% of the failures were below the \ce{CO2}-budget in 2035, only 37\% passed this threshold in 2040. This reminds the importance of near-term, 2025-2030, actions to hope for a successful transition.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Cum_gwp_cost_3.pdf}
\caption{Exploration of the state space over the learning process: distribution of occurrence of cumulative emissions (left) and costs (right). Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. The majority of successful transitions have cumulative emissions much lower than the \ce{CO2}-budget and are cheaper than the REF case. }
\label{fig:Cum_gwp_cost}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{RE_in_mix_Efficiency.pdf}
\caption{Exploration of the state space over the learning process: distribution of occurrence of share of renewable energy carriers in the primary energy mix (left) and efficiency (right).  Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. Integration of local \gls{VRES} at early stages then massive import of electrofuels later are needed to secure successful transitions. Below a near-term threshold ($\sim$60\%), the chances of success are limited, \ie no-go zones. Efficiency is a less valuable information for the agent to succeed transitions as failures and successes indistinguishably spread over the whole range.}
\label{fig:RE_in_mix_Efficiency}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Actions_learning.pdf}
\caption{Exploration of the action space over the learning process: distribution of occurrence of the different actions at the time they are taken by the agent. Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. Besides this wide exploration, successes and failures indistinguishably spread over the whole ranges of actions. In other words, it does not seem to be any clear set of actions to support successful transitions.}
\label{fig:Actions_learning}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Binding_learning.pdf}
\caption{Keeping only the successful transitions, distribution of occurrence of binding and not binding actions. Depending on the action and its timing, it is actually constraining the optimisation through EnergyScope Pathway or not. Sweet spots can be identified when considering the limits of \gls{GWP} and fossil gas consumption. Limiting coal consumption is always constraining, unlike \gls{LFO} that is ``naturally'' substituted by EnergyScope Pathway in the near-term.}
\label{fig:Binding_learning}
\end{figure}

### 2024-07-05 chap_RobPol/chap_RobPol.tex
Rebuttals: Keep going...

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.9\textwidth]{Top5_PC1.pdf}
\includegraphics[width=0.9\textwidth]{Top5_PC2.pdf}
\includegraphics[width=0.9\textwidth]{Top5_PC3.pdf}
\caption{Contribution of the top-5 technologies to the first three \gls{PCs} over the different representative years. Where the variation of \gls{PV} panels supported by the electrification of the \gls{HT} heat via industrial electric resistors is the main varying drivers over the whole transition, there are tipping years like 2025 for the production of \gls{HVC} (\ie naphtha-cracker to \gls{MTO} or 2025-2030 for the production of \gls{LT} heat (\ie from oil and gas boilers to electric \gls{HP}.}
\label{fig:Top5_PC_year}
\end{figure}

>>>>

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.9\textwidth]{Top5_PC1.pdf}
\includegraphics[width=0.9\textwidth]{Top5_PC2.pdf}
\includegraphics[width=0.9\textwidth]{Top5_PC3.pdf}
\caption{Contribution of the top-5 technologies to the first three \gls{PCs} over the different representative years. Where the variation of \gls{PV} panels supported by the electrification of the \gls{HT} heat via industrial electric resistors are the main varying drivers over the whole transition, there are tipping years like 2025 for the production of \gls{HVC} (\ie naphtha-cracker to \gls{MTO}) or 2025-2030 for the production of \gls{LT} heat (\ie from oil and gas boilers to electric \gls{HP}).}
\label{fig:Top5_PC_year}
\end{figure}

### 2024-07-05 chap_atom_mol/chap_atom_mol.tex
Rebuttals: Keep going...

On top of scenarios with more profound behavioural changes, variety of technological pathways are often investigated to meet the ambitions of climate change mitigation. For instance, in their work, \citet{My2050} assessed different scenarios for a climate neutral Belgium by 2050. Depending on the scenarios, the emphasis is put on a higher electrification of the whole-energy system, a higher consumption of hydrogen or more complex moleclules (\ie electrofuels) or a bigger reliance on biomass and the related \gls{BECCS}. Besides these, ``unicorn'' technologies are also investigated \cite{heuberger2018impact}. These technologies have not (yet) reached a high enough maturity, \ie Technology Readiness Level (TRL) 3 to 7, or they face hurdles such as social acceptance. Therefore, they are currently not deployed on a large scale. Among others, nuclear energy finds an interest in the literature \cite{IEA_Nuclear_2022,PATHS2050}. In the form of \acrfull{SMR}, nuclear energy is also discussed in actual current investments (in Belgium for instance \cite{SMRlesoir}).

@@

In their review, \citet{yue2018review} pointed out that uncertainties were accounted for either snapshot models (\ie optimizing a single target future year), to assess a single energy sector (\ie the power sector, most of the time) or with limited number of uncertain parameters, \ie about ten, in a stochastic programming approach to optimise a transition pathway with a small number of time stages, \ie two or three. Here, the uncertainty quantification addresses the total transition pathway of the whole-energy system, with 34 uncertain parameters (see Section \ref{sec:cs:uncertainty}).

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{Elec_Tech_Cap_REF.pdf}
\includegraphics[width=\textwidth]{Elec_Tech_Cap_SMR.pdf}
\caption{As soon as available (\ie 2040), \acrfull{SMR} is deployed to their maximum potential (\ie 6\,GW) to substitute more expensive flexible generation units (\ie gas and ammonia \gls{CCGT}).}
\label{fig:results_deter_tech_cap_elec}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{Elec_Layer_REF.pdf}
\includegraphics[width=\textwidth]{Elec_Layer_SMR.pdf}
\caption{The production of electricity from \gls{SMR} substitutes more efficient technologies (\ie \gls{CCGT} and \gls{CHP}) and boosts the electrification of the rest of the system, mostly the industrial high-temperature heat sector.}
\label{fig:results_deter_layer_elec}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{GWP_per_sector_diff.pdf}
\caption{Emissions per energy sector differerence between the SMR and the REF cases. The deployment of \gls{SMR} from 2040 onward allows maintaining longer the use of \acrfull{LFO} for the production of \acrfull{HVC}.}
\label{fig:GWP_per_sector_diff}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{Primary_mix_REF.pdf}
\includegraphics[width=0.8\textwidth]{Primary_mix_SMR.pdf}
\includegraphics[width=0.8\textwidth]{Primary_mix_diff_uranium.pdf}
\caption{(Top) Primary energy mix over the transition for the REF and SMR cases. (Bottom) Difference of the mix between the two cases after aggregation per category. In line with \citet{rixhon2021terminology},uranium is considered as a non-renewable resource. It is nevertheless dissociated from other non-renewable resources for the sake of clarity. The imported electricity is split between imported non-renewable and renewable resources between the current share of renewables, \ie 37.41\% \cite{eurostat_share_re_elec}, and an assumed 100\%-renewable European electricity mix by 2050.}
\label{fig:results_deter_energy_mix}
\end{figure}

@@

The main impact of including \gls{SMR} from 2040 onward on the high-temperature heat sector is (i) its higher direct electrification and (ii) the reduction of more efficient heat-and-power co-generation in the benefit of single-output industrial boilers. In the REF case, industrial electric heaters are mainly used to absorb, instead of curtailing, the ``over-production" of the 59\,GW solar-PV when fully deployed, in the sunny days. With \gls{SMR}, by 2050, an additional 1.7\,GW (+13\%) of these heaters can rely on a more constant supply of defossilised electricity. This increases the load factor (31\%) and yearly production (48\%) of these industrial electric heaters. Then, given the 44.6\,TWh of electricity produced by \gls{SMR}, cogeneration units are less relevant and, by 2050, 2.6\,GW industrial gas boilers completely substitute \gls{CHP} to produce 16.6\,TWh (\ie 23\%) of the total production of high-temperature heat.\\

@@

\subsection{Total transition cost}
\label{subsec:atom_mol:results_uq_cost}
Exhaustively listed in Appendix \ref{app:UQ_transition_cost}, \Cref{tab:UQ_short} gathers the most impacting parameters on the total transition cost. Per \citet{Turati2017}, parameters are considered as ``impacting'' if their Sobol' index is above the threshold $=1/d$, $d$ being the total number of uncertain parameters after the pre-selection phase. In this case, $d=34$, and, consequently, the threshold is equal to 2.9\%. \Cref{tab:UQ_short} shows that the cost of purchasing electrofuels is the most impacting parameter. On the contrary, the potentiality to install \gls{SMR} and its CAPEX have much lower influence on the variation of the total transition cost. Given the uncertainty characterisation presented in Section \ref{sec:cs:uncertainty}, there are 60\% chance that no \gls{SMR} could be installed. In other words, the variation of the parameter $f_{\mathrm{max,SMR}}$ has zero impact on the variation of the total transition cost in 60\% of the samples. Then, in perspective with the scenario analysis of Section \ref{sec:atom_mol:results_deter}, the 3.3\% reduction has been observed when \gls{SMR} is installed from 2040 onward. This represents only 10\% of the samples. Moreover, when it is possible, \gls{SMR} is always installed thanks to its characteristics: cheap and low-emitting fuel, the long lifetime leading to lower annualised CAPEX and higher salvage value. However, given its limited deployment, up to 6\,GW, \gls{SMR}-related parameters have a lower impact on the variation of the total transition cost. On the contrary, more expensive renewable electrofuels are always imported, to a smaller or larger extent depending on the sample. For instance, in the REF case (Section \ref{subsec:atom_mol:results_deter_overall}), the imported electrofuels represent, by 2050, 152.9\,TWh (\ie 41\% of the primary energy mix) with an average 93€/MWh cost of purchasing and, over the entire transition, a 273\,b€ cumulative OPEX (\ie 25\% of the total transition cost).

@@

\Cref{fig:UQ_PDF_total_transition_cost} shows the \gls{PDF} of the total transition cost given the 1260 samples. Stretching between 660\,b€ and 2050\,b€, the mean, the median and the nominal value (\ie REF case) are close to each other, respectively 1180\,b€, 1160\,b€ and 1080\,b€. Similarly to the analysis carried out by \citet{coppitters2023optimizing}, one observes here that the distribution is right-skewed. It could then be qualified as ``fragile" as the top 50\% of the samples cover a bigger range (\ie between 1160 and 2050) than the bottom 50\% of the samples (\ie between 660 and 1160). In other words, the bad scenarios, resulting in a total transition cost higher than the median, have a bigger effect on this cost.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_PDF_total_transition_cost.pdf}
\caption{\acrfull{PDF} of the total transition cost. The mean, $\mu=1.18\cdot10^3$\,b€, is slightly higher than the median ($P_{50}=1.16\cdot10^3$\,b€) and the nominal cases cost, $1.08\cdot10^3$\,b€ and $1.04\cdot10^3$\,b€ respectively for the REF and SMR cases. Also, with a standard deviation, $\sigma=197$\,b€, a 95\%-confidence interval would be about [0.8; 1.6]$\cdot10^3$\,b€.}
\label{fig:UQ_PDF_total_transition_cost}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\textwidth]{UQ_Electrofuels_no_outlier.pdf}
\includegraphics[height=3.8cm]{Schematic_boxplot_no_outlier.pdf}
\caption{Distribution of the imported renewable electrofuels over the transition. Starting from no electrofuel in 2020, their respective import rises progressively along the transition (\ie increasing median) at different growth rates and with different ranges of values. }
\label{fig:results_uq_electrofuels}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_Methanol_samples_2.pdf}
\caption{Trend lines of the key parameters on the import of e-methanol in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_methanol}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_H2_samples_2.pdf}
\caption{Trend lines of the key parameters on the import of e-hydrogen in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_H2}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_Gas_samples_2.pdf}
\caption{Trend lines of the key parameters on the import of e-methane in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_gas}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_Ammonia_samples_2.pdf}
\caption{Trend lines of the key parameters on the import of e-ammonia in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_ammonia}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_SMR_samples_2.pdf}
\caption{Trend lines of the key parameters on the installed capacity of \gls{SMR} in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_SMR}
\end{figure}

@@

In conclusion, this work puts under the spotlight the ``competition'' between \gls{SMR} and imported electrofuels while both of them support the integration of local \gls{VRES}. Betting on \gls{SMR} means letting the emissions go up at early stages of the transition (\ie \gls{LFO} still used in 2025 to produce \gls{HVC}) to end up with an overall transition and a system by 2050 that are 3.3\% and 8.8\% cheaper than in the REF case, respectively. However, given the need to import molecules at earlier stages of the transition, with or without \gls{SMR}, it seems reasonable to keep on investing in the transport and the infrastructure to support these imports. This would allow covering ourselves against the risk of eventually not having \gls{SMR} available by the end of the first half of this century. The perfect foresight approach does not model this risk as the possibility to install or not \gls{SMR} is known in advance. In the EnergyScope Pathway model, assuming the same technology mix as in the SMR case up to 2040 but without installing \gls{SMR} for the rest of the transition, \ie between 2040 and 2050, would require drastic (and quick) changes to still meet the \ce{CO2}-budget: around 44\,b€ of additional cumulative OPEX in purchasing electrofuels, mostly e-ammonia (+32\,TWh by 2050) and e-methane (+31\,TWh) to supply \gls{CCGT} and industrial \gls{CHP}. As a more concrete example, Fluxys, the manager of the gas network in Belgium, already presents significant investments by 2032 (\ie 1.3\,b€) to support this transition in the near future \cite{Fluxys_2023}.

Chapters \ref{chap:chap_RL} and \ref{chap:chap_RobPol} put in practice the methodology detailed in Chapter \ref{chap:chap_methodo} to come up with a policy to get a system robust to this possibility of finally not having access to \gls{SMR}, as well as other uncertainties, especially when considering a myopic progression through the transition.

>>>>

On top of scenarios with more profound behavioural changes, variety of technological pathways are often investigated to meet the ambitions of climate change mitigation. For instance, in their work, \citet{My2050} assessed different scenarios for a climate neutral Belgium by 2050. Depending on the scenarios, the emphasis is put on a higher electrification of the whole-energy system, a higher consumption of hydrogen or more complex molecules (\ie electrofuels) or a bigger reliance on biomass and the related \gls{BECCS}. Besides these, ``unicorn'' technologies are also investigated \cite{heuberger2018impact}. These technologies have not (yet) reached a high enough maturity, \ie Technology Readiness Level (TRL) 3 to 7, or they face hurdles such as social acceptance. Therefore, they are currently not deployed on a large scale. Among others, nuclear energy finds an interest in the literature \cite{IEA_Nuclear_2022,PATHS2050}. In the form of \acrfull{SMR}, nuclear energy is also discussed in actual current investments (in Belgium for instance \cite{SMRlesoir}).

@@

In their review, \citet{yue2018review} pointed out that uncertainties were accounted for either snapshot models (\ie optimizing a single target future year), to assess a single energy sector (\ie the power sector, most of the time) or with limited number of uncertain parameters, \ie about ten, in a stochastic programming approach to optimise a transition pathway with a small number of time stages, \ie two or three. Here, the uncertainty quantification addresses the total transition pathway of the whole-energy system, with 34 uncertain parameters (see Section \ref{sec:cs:uncertainty}). This \gls{GSA} applied to a model optimising the transition pathway of a whole-energy system with these many uncertain parameters is the main contribution of this chapter. Through this analysis, we also managed to point out the impact of integrating \gls{SMR} in the Belgian energy system as well as the main drivers of the import of e-hydrogen, e-methane, e-ammonia and e-methanol by 2050.

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{Elec_Tech_Cap_REF.pdf}
\includegraphics[width=\textwidth]{Elec_Tech_Cap_SMR.pdf}
\caption{Installed capacities in the power sector in the REF (top) and SMR (down) scenarios. As soon as available (\ie 2040), \acrfull{SMR} is deployed to their maximum potential (\ie 6\,GW) to substitute more expensive flexible generation units (\ie methane and ammonia \gls{CCGT}).}
\label{fig:results_deter_tech_cap_elec}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{Elec_Layer_REF.pdf}
\includegraphics[width=\textwidth]{Elec_Layer_SMR.pdf}
\caption{Balance between the production and the consumption of electricity in the REF (top) and SMR (down) scenarios. The production of electricity from \gls{SMR} substitutes more efficient technologies (\ie \gls{CCGT} and \gls{CHP}) and boosts the electrification of the rest of the system, mostly the industrial high-temperature heat sector.}
\label{fig:results_deter_layer_elec}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=\textwidth]{GWP_per_sector_diff.pdf}
\caption{Emissions per energy sector difference between the SMR and the REF cases. The deployment of \gls{SMR} from 2040 onward allows maintaining longer the use of \acrfull{LFO} for the production of \acrfull{HVC}.}
\label{fig:GWP_per_sector_diff}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{Primary_mix_REF.pdf}
\includegraphics[width=0.8\textwidth]{Primary_mix_SMR.pdf}
\includegraphics[width=0.8\textwidth]{Primary_mix_diff_uranium.pdf}
\caption{(Top) Primary energy mix over the transition for the REF and SMR cases. (Down) Difference of the mix between the two cases after aggregation per category. In line with \citet{rixhon2021terminology},uranium is considered as a non-renewable resource. It is nevertheless dissociated from other non-renewable resources for the sake of clarity. The imported electricity is split between imported non-renewable and renewable resources between the current share of renewables, \ie 37.41\% \cite{eurostat_share_re_elec}, and an assumed 100\%-renewable European electricity mix by 2050.}
\label{fig:results_deter_energy_mix}
\end{figure}

@@

The main impact of including \gls{SMR} from 2040 onward on the high-temperature heat sector is (i) its higher direct electrification and (ii) the reduction of more efficient heat-and-power co-generation in the benefit of single-output industrial boilers. In the REF case, industrial electric heaters are mainly used to absorb, instead of curtailing, the ``over-production'' of the 59\,GW solar-PV when fully deployed, in the sunny days. With \gls{SMR}, by 2050, an additional 1.7\,GW (+13\%) of these heaters can rely on a more constant supply of defossilised electricity. This increases the load factor (31\%) and yearly production (48\%) of these industrial electric heaters. Then, given the 44.6\,TWh of electricity produced by \gls{SMR}, cogeneration units are less relevant and, by 2050, 2.6\,GW industrial gas boilers completely substitute \gls{CHP} to produce 16.6\,TWh (\ie 23\%) of the total production of high-temperature heat.\\

@@

\subsection{Total transition cost}
\label{subsec:atom_mol:results_uq_cost}
Exhaustively listed in Appendix \ref{app:UQ_transition_cost}, \Cref{tab:UQ_short} gathers the most impacting parameters on the total transition cost. Per \citet{Turati2017}, parameters are considered as ``impacting'' if their Sobol' index is above the threshold $=1/d$, $d$ being the total number of uncertain parameters after the pre-selection phase. In this case, $d=34$, and, consequently, the threshold is equal to 2.9\%. \Cref{tab:UQ_short} shows that the cost of purchasing electrofuels is the most impacting parameter. On the contrary, the potentiality to install \gls{SMR} and its CAPEX have much lower influence on the variation of the total transition cost. Given the uncertainty characterisation presented in Section \ref{sec:cs:uncertainty}, there are 60\% chance that no \gls{SMR} could be installed. When $f_{\mathrm{max,SMR}}$ takes a value between 0 and 0.59, the variation of the total transition cost is only due to the variation of the other parameters. Then, in perspective with the scenario analysis of Section \ref{sec:atom_mol:results_deter}, the 3.3\% reduction has been observed when \gls{SMR} is installed from 2040 onward. This represents only 10\% of the samples. Moreover, when it is possible, \gls{SMR} is always installed thanks to its characteristics: cheap and low-emitting fuel, the long lifetime leading to lower annualised CAPEX and higher salvage value. However, given its limited deployment, up to 6\,GW, \gls{SMR}-related parameters have a lower impact on the variation of the total transition cost. On the contrary, more expensive renewable electrofuels are always imported, to a smaller or larger extent depending on the sample. For instance, in the REF case (Section \ref{subsec:atom_mol:results_deter_overall}), the imported electrofuels represent, by 2050, 152.9\,TWh (\ie 41\% of the primary energy mix) with an average 93€/MWh cost of purchasing and, over the entire transition, a 273\,b€ cumulative OPEX (\ie 25\% of the total transition cost).

@@

\Cref{fig:UQ_PDF_total_transition_cost} shows the distribution of the total transition cost given the 1260 samples. Stretching between 660\,b€ and 2050\,b€, the mean, the median and the nominal value (\ie REF case) are close to each other, respectively 1180\,b€, 1160\,b€ and 1080\,b€. Similarly to the analysis carried out by \citet{coppitters2023optimizing}, one observes here that the distribution is right-skewed. It could then be qualified as ``fragile'' as the top 50\% of the samples cover a bigger range (\ie between 1160 and 2050) than the bottom 50\% of the samples (\ie between 660 and 1160). In other words, the bad scenarios, resulting in a total transition cost higher than the median, have a bigger effect on this cost.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_PDF_total_transition_cost.pdf}
\caption{Distribution of the total transition cost over the 1260 runs of the \acrfull{GSA}. The mean, $\mu=1.18\cdot10^3$\,b€, is slightly higher than the median ($P_{50}=1.16\cdot10^3$\,b€) and the nominal cases cost, $1.08\cdot10^3$\,b€ and $1.04\cdot10^3$\,b€ respectively for the REF and SMR cases. Also, with a standard deviation, $\sigma=197$\,b€, a 95\%-confidence interval would be about [0.8; 1.6]$\cdot10^3$\,b€.}
\label{fig:UQ_PDF_total_transition_cost}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\textwidth]{UQ_Electrofuels_no_outlier.pdf}
\includegraphics[height=3.8cm]{Schematic_boxplot_no_outlier.pdf}
\caption{Distribution of the imported renewable electrofuels over the transition. Starting from no electrofuel in 2020, their respective import rises progressively along the transition at different growth rates and with different ranges of values. }
\label{fig:results_uq_electrofuels}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_Methanol_samples_2.pdf}
\caption{Trend lines of the key parameters (and their Sobol' index) on the import of e-methanol in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_methanol}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_H2_samples_2.pdf}
\caption{Trend lines of the key parameters (and their Sobol' index) on the import of e-hydrogen in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_H2}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_Gas_samples_2.pdf}
\caption{Trend lines of the key parameters (and their Sobol' index) on the import of e-methane in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_gas}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_Ammonia_samples_2.pdf}
\caption{Trend lines of the key parameters (and their Sobol' index) on the import of e-ammonia in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_ammonia}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{UQ_SMR_samples_2.pdf}
\caption{Trend lines of the key parameters (and their Sobol' index) on the installed capacity of \gls{SMR} in 2050. Around these lines, box plots point out the distribution of the output of interest for the extreme values (either bottom-15\% or top-15\%) of some parameters. The grey dashed line gives the value of the output of interest in the REF case. }
\label{fig:results_uq_samples_SMR}
\end{figure}

@@

In conclusion, this work puts under the spotlight the ``competition'' between \gls{SMR} and imported electrofuels while both of them support the integration of local \gls{VRES}. Betting on \gls{SMR} means letting the emissions go up at early stages of the transition (\ie \gls{LFO} still used in 2025 to produce \gls{HVC}) to end up with an overall transition and a system by 2050 that are 3.3\% and 8.8\% cheaper than in the REF case, respectively. However, given the need to import molecules at earlier stages of the transition, with or without \gls{SMR}, it seems reasonable to keep on investing in the transport and the infrastructure to support these imports. In practice, Fluxys, the manager of the gas network in Belgium, already presents significant investments by 2032 (\ie 1.3\,b€) to support this transition in the near future \cite{Fluxys_2023}. The investments into the molecules would allow covering ourselves against the risk of eventually not having \gls{SMR} available by the end of the first half of this century. The perfect foresight approach does not model this risk as the possibility to install or not \gls{SMR} is known in advance. In the EnergyScope Pathway model, assuming the same technology mix as in the SMR case up to 2040 but without installing \gls{SMR} for the rest of the transition, \ie between 2040 and 2050, would require drastic (and quick) changes to still meet the \ce{CO2}-budget: around 44\,b€ of additional cumulative OPEX in purchasing electrofuels, mostly e-ammonia (+32\,TWh by 2050) and e-methane (+31\,TWh) to supply \gls{CCGT} and industrial \gls{CHP}. 
%
%Chapters \ref{chap:chap_RL} and \ref{chap:chap_RobPol} put in practice the methodology detailed in Chapter \ref{chap:chap_methodo} to come up with a policy to get a system robust to this possibility of finally not having access to \gls{SMR}, as well as other uncertainties, especially when considering a myopic progression through the transition.
